\pkg{fitdistrplus}: An \proglang{R} Package for Distribution Fitting Methods

### 2013-04-03 JSS/fitdistrplus/paper2JSS.Rnw
work on the bibtex files
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/riskassessment@241 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

Maximum likelihood and moment matching 
%(the default on many softwares)
% si on ne trouve pas de references on l'enleve?
estimations are 
certainly the most commonly used method for fitting distributions. 
Keeping in mind that these two methods may produce very different results,
the user should be aware of its great sensitivity to outliers
when choosing the matching moment estimation (MME). This may be seen as an advantage in our example if the
objective is to better describe the right tail of the distribution, but it may be seen as a
drawback if the objective is different.
% As-tu des references indiquant que le MLE est meilleurs dans certains cas, cf. commentaires de Regis ?

%\subsubsection{Quantile matching estimation}
%\label{QME}
%%%%%%%%%%
Fitting of a parametric distribution may also be done by matching theoretical quantiles
of the parametric distributions (for specified probabilities) against the empirical quantiles.
Equation (\ref{quantiles}) below is thus very similar to Equations (\ref{moments}) and (\ref{centmoments})
  \begin{equation}
    \label{quantiles}
    F^{-1}(p_{k}|\theta)=Q_{n,p_{k}} 
  \end{equation}
for $k=1,\ldots,d$, with $d$ the number of parameters to estimate (dimension of $\theta$ if there is
no fixed parameters) and
 $Q_{n,p_{k}}$ the empirical quantiles calculated from data for specified probabilities $p_{k}$.

@@

The \proglang{R}  package \pkg{fitdistrplus} allows to easily fit distributions. 
Our main objective while developing this package was to provide tools for helping 
\proglang{R} users to fit distributions to data. 
We have been encouraged to pursue our work by feedbacks 
from users of our package in various areas as food or environmental risk assessment, epidemiology, ecology, 
molecular biology, genomics, bioinformatics, hydraulics, mechanics, financial and actuarial mathematics 
or operations research. Indeed, this package is already used a lot by practionners and academics 
for simple MLE fits in \cite{distrModJSS,jaloustreetal11,saketal11,wilson12,kochetal12,marquetouxetal12,nordan12,
scholletal12,suuronenetal12,varoetal12,mandletal13}, for MLE fits and goodness-of-fit statistics in 
\cite{garcia12,tarnczi11,bagariaetal12,benavidesetal12,breitbach12,poduje12,pouillotetal10}, for MLE fits 
and bootstrap in  \cite{croucheretal12,meheustetal12,orellanoetal12,telloetal12,hoelzeretal12}, for MLE fits, 
bootstrap and goodness-of-fit statistics in \cite{rebuge12}, for MME fit in \cite{luangkesornetal12}, for 
censored MLE and bootstrap in \cite{lehaetal11,poulliotetal12,jongenburgeretal12,commeauetal12,pouillotetal11}, 
for graphic analysing in \cite{anandetal12,rosa12}, for grouped-data fitting methods in  \cite{fusteinercostafreda12} 
and more generally in  \cite{eling12,busschaertetal10,brooksetal11,aktassjostrand11,eikhermann12,bakos11}.
Many extensions of this package are planned in the future: 
we target to extend to censored data some methods for the moment only available for non-censored data, 
especially concerning goodness-of-fit evaluation and fitting methods. We will also enlarge the choice of 
fitting methods for non-censored data, by proposing new goodness-of-fit distances (e.g. distances 
based on quantiles) for maximum goodness-of-fit estimation 
and new types of moments (e.g. limited expected values) for moment matching estimation. At last, we will
consider the case of multivariate distribution fitting.

>>>>

Maximum likelihood and moment matching 
%(the default on many softwares)
% si on ne trouve pas de references on l'enleve?
estimations are 
certainly the most commonly used method for fitting distributions (\cite{Cullen99}). 
Keeping in mind that these two methods may produce very different results,
the user should be aware of its great sensitivity to outliers
when choosing the matching moment estimation (MME). This may be seen as an advantage in our example if the
objective is to better describe the right tail of the distribution, but it may be seen as a
drawback if the objective is different.
% As-tu des references indiquant que le MLE est meilleurs dans certains cas, cf. commentaires de Regis ?

%\subsubsection{Quantile matching estimation}
%\label{QME}
%%%%%%%%%%
Fitting of a parametric distribution may also be done by matching theoretical quantiles
of the parametric distributions (for specified probabilities) against the empirical quantiles (\cite{Tse2009}).
Equation (\ref{quantiles}) below is thus very similar to Equations (\ref{moments}) and (\ref{centmoments})
  \begin{equation}
    \label{quantiles}
    F^{-1}(p_{k}|\theta)=Q_{n,p_{k}} 
  \end{equation}
for $k=1,\ldots,d$, with $d$ the number of parameters to estimate (dimension of $\theta$ if there is
no fixed parameters) and
 $Q_{n,p_{k}}$ the empirical quantiles calculated from data for specified probabilities $p_{k}$.

@@

The \proglang{R}  package \pkg{fitdistrplus} allows to easily fit distributions. 
Our main objective while developing this package was to provide tools for helping 
\proglang{R} users to fit distributions to data. 
We have been encouraged to pursue our work by feedbacks 
from users of our package in various areas as food or environmental risk assessment, epidemiology, ecology, 
molecular biology, genomics, bioinformatics, hydraulics, mechanics, financial and actuarial mathematics 
or operations research. Indeed, this package is already used a lot by practionners and academics 
for simple MLE fits in \cite{distrModJSS,jaloustreetal11,saketal11,kochetal12,marquetouxetal12,nordan12,
scholletal12,suuronenetal12,varoetal12,mandletal13}, for MLE fits and goodness-of-fit statistics in 
\cite{garcia12,tarnczi11,bagariaetal12,benavidesetal12,breitbach12,poduje12,Pouillot10}, for MLE fits 
and bootstrap in  \cite{croucheretal12,meheustetal12,orellanoetal12,telloetal12,hoelzeretal12}, for MLE fits, 
bootstrap and goodness-of-fit statistics in \cite{rebuge12}, for MME fit in \cite{luangkesornetal12}, for 
censored MLE and bootstrap in \cite{lehaetal11,poulliotetal12,jongenburgeretal12,commeauetal12,pouillotetal11}, 
for graphic analysing in \cite{anandetal12,rosa12}, for grouped-data fitting methods in  \cite{fusteinercostafreda12} 
and more generally in  \cite{eling12,busschaertetal10,brooksetal11,aktassjostrand11,eikhermann12,bakos11}.
Many extensions of this package are planned in the future: 
we target to extend to censored data some methods for the moment only available for non-censored data, 
especially concerning goodness-of-fit evaluation and fitting methods. We will also enlarge the choice of 
fitting methods for non-censored data, by proposing new goodness-of-fit distances (e.g. distances 
based on quantiles) for maximum goodness-of-fit estimation 
and new types of moments (e.g. limited expected values) for moment matching estimation. At last, we will
consider the case of multivariate distribution fitting.
